---
title: "PSTAT 174 HW 4 R Code"
author: "Idris Rasheed"
date: "May 4, 2017"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

###Problem 4
(a) Plot the time series and find its mean, variance, acf and pacf. Notice the presence of a strong seasonal component with period 12 in the graphs of the data and in the sample acf. Comment on all plots.
```{r Series Plot}
#Opens tools from tseries package
library(tseries)
#reads in deaths dataset
deaths = read.table("deaths.txt", sep = ",")
#Plot time series for deaths
ts.plot(deaths)
title(expression(paste("Series")), line = -1, outer=TRUE)
```

```{r ACF and PACF}
#Makes deaths txt a time series object
deaths_ts <- ts(deaths)
#combines plots
graph <- par(mfrow = c(1,2))
#Plots ACF for deaths data
acf(deaths_ts)
#Plots PACF
pacf(deaths_ts)
title(expression(paste("Deaths")), line = -1, outer=TRUE)
#Plots ACF and PACG sibe by side
par(graph)
```

```{r Mean}
#computes Mean
mean(deaths_ts)
```

```{r Variance}
#Computes Variance 
var(deaths)
```

(b) Use differencing to deseasonalize and detrend the data. On each step include graphs of the
data, corresponding variance, acf and pacf, and comment on your decisions to difference (or not)
and at what lag.

```{r differencing 1}
#Differencing at lag 1
dif_lag_one <- diff(deaths_ts, lag = 1)
ts.plot(dif_lag_one ,main = "Differenced Deaths at lag = 1")
```

```{r Var of Dif 1}
var(dif_lag_one)
```

```{r differencing 1}
#combines plots
graph <- par(mfrow = c(1,2))
#Plots ACF for difference at lag 1
acf(dif_lag_one)
#Plots PACF for difference at lag 1
pacf(dif_lag_one)
title(expression(paste("Difference at Lag 1")), line = -1, outer=TRUE)
#Plots ACF and PACG sibe by side
par(graph)
```

```{r differencing 2}
#Differencing at lag 2
dif_lag_two <- diff(deaths_ts, lag = 2)
ts.plot(dif_lag_two ,main = "Differenced Deaths at lag = 2")
```

```{r Var of Dif 2}
var(dif_lag_two)
```

```{r differencing 2}
#combines plots
graph <- par(mfrow = c(1,2))
#Plots ACF for difference at lag 1
acf(dif_lag_two)
#Plots PACF for difference at lag 1
pacf(dif_lag_two)
title(expression(paste("Difference at Lag 2")), line = -1, outer=TRUE)
#Plots ACF and PACG sibe by side
par(graph)
```
